---
title: "Pokemon EDA"
output: 
  html_document:
    toc: true
    toc_depth: 2
editor_options: 
  chunk_output_type: console
---

```{r Importing Libraries and Data, warning=F, message=F}

if (!require(pacman)){
  install.packages("pacman")
  library(pacman)
}

p_load(httr, jsonlite, rvest, xml2, readr, here, diffdf,
       tidyr, ggplot2, dplyr, tibble, magrittr, purrr)

WD <- getwd()
pokedata <- read_csv("pokedata/pokemon.csv")
```

```{r JSON, warning=FALSE, include=F}

# The response 'res' an object obtained by using the GET function on a link.
# res <- GET("https://pokemondb.net/pokedex/all")
# pokedex <- fromJSON(rawToChar(res$content))


# NOTE: WEBSITE DOES NOT HAVE AN API (JSON).
# PIVOT TO USING HTML WEB SCRAPING.
```

Initial findings show that [PokemonDB.net](https://pokemondb.net/pokedex/all) does not have an API we can use to extract data in JSON format; one thing to note is that the the site displays a table of all the Pokemon, with important data points such as ID, name, typing, and its in-game stats. Because the website has a formatted table, the *Inspect Element* tool that most browsers have came in handy, where I was able to determine the HTML node that contains the table with the desired data.

As I worked through the data cleaning and wrangling steps, I realized that I could actually compile a 7 that included more information such as which Generation a Pokemon was introduced in, and whether or not the Pokemon was a Legendary. To accomplish this, I had to scrape some more data from other websites, to which [Bulbapedia](https://bulbapedia.bulbagarden.net) came to mind first, as it was a resource that I commonly used when I play through the mainline games.

A brief description on the data I'm scraping:

-   The National Pokedex, *pdex*, which is a list of all existing Pokemon in the mainline games. This data was sourced from PokemonDB.net. The data does not include which generation it was from and does not indicate whether or not a Pokemon was a Legendary.
-   The Mega-Evolutions, *gen6Megas* and *gen7*Megas, which lists all of the existing Mega-Evolution capable Pokemon. This is an exhaustive list as these Pokemon forms were introduced only in Generations 6 and 7, so moving forward is a non-issue (thus far).
-   The Generation, *natdex*, which lists each Pokemon by their respective generation in which they were introduced. This is comprehensive as of the release of *Pokemon Sword* and *Pokemon Shield*, including all of the expansion passes.

General notes on scraping method:

-   The approach I took was somewhat tedious; I had to first find the website with the data I'm looking for, then I used the 'Inspect Element' tool and looked for the specific HTML Node, the specific node paths can be copied and pasted into the code.

    -   For *pdex*, it was simple because there was only a single table on the web page, so all I had to do was just type *html_node("table")* and pipe it into the table fill function.

    -   The *gen6Megas* and *gen7Megas* were a bit more complicated as I had to specify which tables to grab. Fortunately, I was able to find that it had a specific class in the table called 'roundy' and that I was able to specify something called an 'xpath' to retrieve this data. From here I was able to pipe the output to retrieve the tables for each generation of Mega-Evolution (*.[[1]]* for Generation 6, *.[[2]]* for Generation 7).

    -   Now *natdex* was particularly difficult because the HTML Node for the table did not have specific class name, so I had to manually figure out the xpath. Note that the index for the data of interest for the tables start at [[2]] and go to [[9]] (Generations 1 through 8).

-   I made the scraped data variables hidden in case I have to rerun the wrangling code while making any quick adjustments. This way, I won't get an error since the wrangled data gets written into a new variable.

-   While I am positive there is a more efficient way to get the Generation data from the HTML tables, I found that it drove me less crazy to just hard-code the tables I want (since it is a small, exhaustive list).

```{r Webscraping Data with HTML}
.pdex <- read_html("https://pokemondb.net/pokedex/all")

.megas <- read_html("https://bulbapedia.bulbagarden.net/wiki/Mega_Evolution")

.natdex <- read_html("https://bulbapedia.bulbagarden.net/wiki/List_of_Pokémon_by_National_Pokédex_number")

.pdex <- .pdex %>% 
  html_node("table") %>% 
  html_table(fill = T)

.megas_list <- .megas %>% 
  html_nodes(xpath = ".//table[contains(@class, 'roundy')]") %>% 
  html_table(fill = T)

  .gen6Megas <- .megas_list %>% 
    .[[1]] %>% 
    .[-1, c(-2, -5)]
  
  .gen7Megas <- .megas_list %>% 
    .[[2]] %>% 
    .[-1, c(-2, -5)]

.gen1All <- .natdex %>% 
  html_nodes(xpath = "//*[@id='mw-content-text']/table[2]") %>% 
  html_table(fill = T) %>% 
  .[[1]] %>% .[, -3]

.gen2All <- .natdex %>% 
  html_nodes(xpath = "//*[@id='mw-content-text']/table") %>%
  html_table(fill =T) %>% 
  .[[3]] %>% .[, -3]

.gen_list <- .natdex %>% 
  html_nodes(xpath = "//*[@id='mw-content-text']/table") %>% 
  html_table(fill = T)

  .gen1All <- .gen_list %>% 
    .[[2]] %>% .[, -3]
  
  .gen2All <- .gen_list %>% 
    .[[3]] %>% .[, -3]

# .gen_list %>% .[[2:10]] %>% View()
# .gen_list_clean <- .gen_list[[2:10]]

  
.pdex %>% 
  html_node("table") %>% 
  html_table(fill = T)

.gen_list_clean <- .gen_list$.[[2:10]]
  
//*[@id="mw-content-text"]/table[2]
//*[@id="mw-content-text"]/table[3]
//*[@id="mw-content-text"]/table[9]
```

```{r}
///
def genList(N)
  .genNlist <- .gen_list %>% 
    .[[N]] %>% .[, -3]
  return genNlist

.gen1All = genNlist(2)
///
```

The data I scraped from these websites has a few things that I needed to adjust for analysis:

-   The Type variable in the Pokedex set contains both a Primary and a Secondary type of a Pokemon. While some Pokemon do not have a secondary typing, we need to split the column into two columns as distinct Types.

-   The columns in the Pokedex set containing different stats have special characters such as '.' and spaces, which we will need to rename. Basic tidying work, but it is simple and helps reduce future keystrokes.

-   The ID numbers in the Pokedex set are identical for Pokemon with multiple forms (e.g., mega-evolutions, regional variants, environmental forms). We will need to assign each Pokemon with unique ID / primary key. Note: The PokeDex number is the same regardless of form.

-   Certain Pokemon have multiple forms, as mentioned above regarding the ID numbers, as such we need to account for distinct form (e.g., mega-evolutions, regional variants, etc.). I use basic regular expression matching to determine where in the Name string to split the columns by.

-   Since the Pokemon were introduced across multiple generations, I had to assign each Pokemon to their respective generations --- this was made complicated by the fact that many Pokemon have multiple forms (as previously mentioned, e.g., mega-evolutions, regional variants, etc.) so I implemented a vectorised *if_else()* statement with *dplyr*.

```{r Cleaning and Wrangling the Pokedex Dataset}

pokedex <- .pdex %>% 
  as_tibble() %>% 
  separate(Type, into = c("PrimaryType", "SecondaryType"),
           sep = "(?<=[a-z])(?=[A-Z])", fill = "right") %>% 
  separate(Name, into = c("Name", "Form"),
           sep = "(?<=[a-z]|[15]0)(?=[15A-Z])", fill = "right") %>%
  rename(DexNo = `#`, Att = Attack, Def = Defense, 
         SpAtt = `Sp. Atk`, SpDef = `Sp. Def`, Spd = Speed) %>% 
  mutate(ID = seq(1:nrow(.pdex))) %>% 
  mutate(Form = replace_na(Form, "Base")) %>% 
  select(ID, DexNo:Spd) 

pokedex %>% 
  mutate(Generation = case_when(
    grepl(paste(.gen6Megas[,1], collapse = "|"), Form) ~ 6,
    grepl(paste(.gen7Megas[,1], collapse = "|"), Form) ~ 7,
    grepl("Alola", Form) ~ 7,
    grepl("Galar", Form) ~ 8,
  )) %>% View()

pokedex[sample(nrow(pokedex), 15),]
```

Some things I learned:

1.  Honestly, these functions by themselves are neither difficult nor complex by themselves, it just looks more complicated because it is chained together, making it look more complicated than it actually is. I chained the functions together to save some screen space.

2.  Regular expressions are (still) confusing.

3.  Regarding the *mutate()* function used to assign a Pokemon's Generation, I used the *dplyr::case_when()* and *grepl()* functions to assign the specific generation values to each Pokemon. The *case_when()* function allowed me to efficiently assign each Pokemon based on certain information, such as if it had specific forms or not.

4.  The *grepl()* function seems to work only with one pattern at a time (vector of length 1), so for multiple patterns that needs matching, I used the separator, *collapse = "\|"*, to basically combine the pattern into a single *grep* pattern (i.e., "\|" acts as an "or" statement for our regular expression, so it **technically** counts as a single pattern).

> Personal Comments: Figuring out the separation regular expression, *sep = "(?\<=[a-z]\|[15]0)(?=[15A-Z])"*, took a lot of trial and error --- solid 30 minutes to find something that reliably works. Typically, you would be able to separate using the regular expression (regex) used for separating *Name* and *Form*, but because of the fact that Zygarde had a *10% Forme* and a *50% Forme*, coupled with *Porygon2* being a distinct Pokemon, you can't simply just include a **[0-9]** match for the look-behind or the look-ahead. Unfortunately, this meant that I had to hard-code **\|[15]0** and **15** into the look-behind and look-ahead.
>
> Of course, this means that this regex would absolutely fail if GameFreak decided to add Pokemon forms that had the specific number '2', similar to how *Porygon2* is named. Sure, I could have just manually edited the data, but that would be no fun and practicing regex is probably important.

Below is a comparison of the two different data sets that I will be using. The first is *pdex*, which is the set that we extracted from the PokemonDB website, the second is *pokedata*, which is the set that was downloaded from Kaggle. The primary differences one will note are:

-   Set *pdex* is a comprehensive list of all existing Pokemon, ranging from Generation 1 through 8, whereas set *pokedata* is a list of Pokemon from Generation 1 through 6.

-   Notably, *pokedata* contains information not in *pdex*. For example, *pokedata* specifies the Generation that a Pokemon debuted in, as well as a True/False indicator regarding a Pokemon's Legendary status.

Before true comparisons are made, it is important that we make sure the formats of both datasets are consistent with each other, else the comparisons won't be as meaningful.

```{r Cleaning and Wrangling the Pokedata Dataset}

pokedata <- pokedata %>% 
  as_tibble() %>% 
  rename(ID = `#`, PrimaryType = `Type 1`,SecondaryType = `Type 2`, 
                    Att = Attack, Def = Defense, 
                    SpAtt = `Sp. Atk`, SpDef = `Sp. Def`, Spd = Speed) %>% 
  mutate(Total = (HP + Att + Def + SpAtt + SpDef + Spd)) %>% 
  select(ID:SecondaryType, Total, HP:Legendary)


```

```{r}


```
